---
title: "Descriptives by Group"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
params:
  num_clusters: 3
  survey_data: "survey_likert"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readr)
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(RColorBrewer)
library(knitr)
library(kableExtra)

source(here("./R/plot_question_bar.R"))
source(here("./R/likert.R"))
source(here("./R/recode_occupation.R"))
source(here::here("./R/save_analysis_edt.R"))

# params <- list(); params$num_clusters <- 4; params$survey_data <- "survey_likert"

print(params)

fs::dir_create(here("./output", "persona",
                    glue::glue("{params$survey_data}"),
                    glue::glue("group_descriptives_{params$num_clusters}")), recurse = TRUE)

# generated from clustering.Rmd
persona <- readr::read_tsv(here(glue::glue("./data/final/persona/{params$survey_data}/persona_group_{params$num_clusters}.tsv")))
metadata <- jsonlite::read_json(here("./data/final/01-surveys/01-self_assessment_persona_metadata.json"))

gp_col <- glue::glue("group{params$num_clusters}")

# i'm using !!sym because we're not in a function
counts <- persona %>%
  dplyr::group_by((!!sym(gp_col)), qbase, question_text) %>%
  dplyr::count(response) %>%
  tidyr::drop_na() %>%
  dplyr::ungroup()
```

```{r}
persona$id_person %>% unique() %>% length() # 67
```


```{r}
persona %>% filter(!is.na(group3)) %>% distinct(id_person) %>% nrow() # 57
```


```{r}
knitr::include_graphics(here::here(glue::glue(
  "./output/persona/{params$survey_data}/dendogram_{params$num_clusters}.png"
)))
```

```{r}
knitr::include_graphics(here::here(glue::glue(
  "./output/persona/{params$survey_data}/dendogram_{params$num_clusters}_gg.png"
)))
```

Missing groups come from data used for validation keeping complete cases.

```{r}
persona %>%
  dplyr::filter(qbase == "Q1.3") %>%
  dplyr::count( (!!sym(gp_col)) )
```

```{r}
occupation_cluster <- persona %>%
  dplyr::filter(qbase == "Q2.3") %>%
  dplyr::select(id_person, id_response, question_part, response, qbase, qid, (!!sym(gp_col))) %>%
  .GlobalEnv$recode_occupation(response) %>%
  tidyr::drop_na()
```

```{r}
g <- ggplot(occupation_cluster, aes(x = occupation_group,
                               fill = as.factor((!!sym(gp_col))) )) +
  geom_bar(position = "dodge") +
  scale_fill_discrete(name = "Cluster") +
  xlab("Occupation Group") +
  ylab("Count") +
  coord_flip() +
  theme_minimal(base_size = 15) +
  theme(text = element_text(family = "Atkinson Hyperlegible"),)
g
```

```{r}
pth <- here("./output",
            "persona",
            glue::glue("{params$survey_data}"),
            glue::glue("group_descriptives_{params$num_clusters}"),
            glue::glue("group_cluster_occupation.png"))

if (params$num_clusters == 3) {
  save_analysis_edt(
    g = g,
    analysis_pth = pth,
    edt_pth = here::here(glue::glue("../dissertation-edt/figs/020-persona/{params$survey_data}/group_cluster_occupation_{params$num_clusters}.png")),
    bg = "white"
  )
} else {
  ggsave(filename = pth, plot = g, width = 8, height = 5, units = "in", bg = "white")
}
```

```{r}
# save out data just for clinical subanalysis
clinician_id_person <- occupation_cluster %>%
  dplyr::filter(occupation_group == "clinician") %>%
  pull(id_person) %>%
  unique()

clinician_dat <- persona %>%
  dplyr::filter(id_person %in% clinician_id_person)

readr::write_tsv(clinician_dat,
                 here(glue::glue("./data/final/persona/{params$survey_data}/persona_clinician_group_{params$num_clusters}.tsv")))
```

# Question Bar Graphs

```{r}
brewer.pal(8, "Dark2")
display.brewer.pal(8, "Dark2")
COLORS <- brewer.pal(8, "Dark2")[c(1:4, 6)]
```

## Main plots

```{r}
for (q in unique(counts$qbase)) {
  print(q)
  #browser()
  if (q %in% c("Q7.2", "Q7.3", "Q7.4")) {next}

  g <- plot_question_bar(dplyr::filter(counts, qbase == q), fill_var = gp_col)
  print(g)
  
  pth <- here("./output",
              "persona",
              glue::glue("{params$survey_data}"),
              glue::glue("group_descriptives_{params$num_clusters}"),
              glue::glue("{q}.png"))
  print(pth)
  
  if (q %in% c("Q2.3", "Q3.4", "Q4.2")) { # plots used for dissertation
    edt_pth <- here("..",
              "dissertation-edt",
              "figs",
              "020-persona",
              glue::glue("{params$survey_data}"),
              glue::glue("{q}-group-{params$num_clusters}.png"))
    save_analysis_edt(
      g = g,
      analysis_pth = pth,
      edt_pth = edt_pth,
      width = 8, height = 5, units = "in", bg = "white"
    )
  } else {
    ggsave(filename = pth, plot = g, width = 8, height = 5, units = "in", bg = "white")
  }
}
```

## Dissertation plots

```{r}
# hacky stuff to just get figs for defense
# Q3.4, Q7.2_7, and Q4.2 had the highest loadings
#6.2, 4.4, and 3.3 were wave 1 questions
specific_plots_ordered <- c("Q3.4", "Q4.2", "Q6.2", "Q4.4", "Q3.3")

counts_defense <- counts %>%
  dplyr::filter(qbase %in% specific_plots_ordered)
  
recode_text <- readr::read_tsv(here::here("data/original/surveys_recodes/question_text_recodes.tsv"))
recode_type <- readr::read_csv(here::here("data/original/surveys_recodes/survey_question_recodes_type.csv")) %>%
  dplyr::filter(survey == "persona")

recode_text <- recode_type %>%
  dplyr::left_join(recode_text, by = "scale_type")

title_wrap_len = 60
question_text_len = title_wrap_len * 2

counts_defense_fixed <- counts_defense %>%
  dplyr::mutate(response = stringr::str_remove_all(response, "\\."),
                response = stringr::str_replace_all(response, "’", "'")) %>%
  dplyr::left_join(recode_text, by = c("qbase" = "question", "response" = "text")) %>%
  dplyr::mutate(response = stringr::str_wrap(response, 30),
                question_text = stringr::str_replace_all(question_text, '\"\"\"\"', '"'),
                question_text = stringr::str_trunc(question_text, width = question_text_len)
                )

for (q in specific_plots_ordered) {
  print(q)
  # q <- specific_plots_ordered[[2]]
  dat <- dplyr::filter(counts_defense_fixed, qbase == q)
  g <- dat %>%
    ggplot(aes(x = forcats::fct_reorder(response, scale_value),
               y = n,
               fill = as.factor(group3))) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_discrete(name = "Group") +
    scale_x_discrete() +
    xlab("") +
    ylab("Count") +
    #labels(line_break(dat$response)) +
    coord_flip() +
    ggtitle(stringr::str_wrap(dat$question_text, title_wrap_len)) +
    theme_minimal(base_size = 15) +
    theme(
      text = element_text(family = "Atkinson Hyperlegible"),
      plot.title.position = "plot",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  print(g)
  
  pth <- here("./output",
              "persona",
              glue::glue("{params$survey_data}"),
              glue::glue("group_descriptives_{params$num_clusters}"),
              glue::glue("{q}.png"))
  
  edt_pth <- here("..",
                  "dissertation-edt",
                  "figs",
                  "020-persona",
                  glue::glue("{params$survey_data}"),
                  glue::glue("{q}-group-{params$num_clusters}.png"))
  save_analysis_edt(
    g = g,
    analysis_pth = pth,
    edt_pth = edt_pth,
    width = 8, height = 5, units = "in", bg = "white"
  )
}
```

# Likert table

```{r, fig.width=8, fig.height=5}
qbase <- "Q7.2"
qid <- get_qid_from_qbase(persona, qbase)
qid

likert_q_count <- likert_calculation(persona, metadata, qbase, qid, grp_var = gp_col)
g <- likert_plot(likert_q_count %>% tidyr::drop_na(), facet_1 = gp_col)
g
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("group_descriptives_{params$num_clusters}"),
                       "likert.png"),
       plot = g,
       width = 13,
       height = 8,
       units = "in")
```

## Separate plots for likert questions

```{r}
dat <- likert_q_count %>%
  tidyr::drop_na() %>%
  dplyr::filter(text == "Using a programming language (like R or Python) can make my analyses easier to reproduce.") %>%
  dplyr::left_join(recode_text %>%
                     filter(question == "Q7.2_7"),
                   by = c("response" = "text"))

g <- dat %>%
  ggplot(aes(x = forcats::fct_reorder(response, scale_value), y = n, fill = as.factor(group3))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  ggtitle(stringr::str_wrap("Using a programming language (like R or Python) can make my analyses easier to reproduce.", 60)) +
  xlab("") +
  ylab("count") +
  scale_fill_discrete("Group") +
  theme_minimal(base_size = 15) +
  theme(
    text = element_text(family = "Atkinson Hyperlegible"),
    plot.title.position = "plot"
  )

print(g)

analysis_pth <- here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("group_descriptives_{params$num_clusters}"),
                       "Q7.2_7.png")

edt_pth <- here("../",
                "dissertation-edt",
                "figs",
                "020-persona",
                glue::glue("{params$survey_data}"),
                glue::glue("Q7.2_7-{params$num_clusters}.png")
                )

save_analysis_edt(
  g = g,
  analysis_pth = analysis_pth,
  edt_pth = edt_pth,
  width = 8, height = 5, units = "in", bg = "white")
  
```


## Heatmap by proportion in group

```{r}
likert_q_count_n_prop <- likert_q_count %>%
  dplyr::mutate(rel_freq = n / sum(n),
                pct = round(rel_freq * 100))
```

```{r, fig.width=13, fig.height=8}
g <- likert_plot(likert_q_count_n_prop %>% tidyr::drop_na(), count_col = "pct", facet_1 = gp_col)
g
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("group_descriptives_{params$num_clusters}"),
                       "likert_prop.png"),
       plot = g,
       width = 13,
       height = 8,
       units = "in")
```

```{r}
# likert plot for poster
poster <- likert_plot(likert_q_count_n_prop %>% tidyr::drop_na(),
                      count_col = "pct") +
  facet_wrap(as.formula(glue::glue("~ {gp_col}")), ncol = 1) +
  theme(legend.position = "none",
        text = element_text(size=12)) +
  ggtitle("Percent of Responses per Group")
poster
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("group_descriptives_{params$num_clusters}"),
                       "likert_prop_poster.png"),
       plot = poster,
       width = 8,
       height = 13,
       units = "in")
```


# Free Response

- Q7.3: Please share what you most hope to learn from participating in this workshop and/or workshop series. 
- Q7.4: What do you want to know or be able to do after this workshop (or series of sessions) that you don't know or can't do right now?

## Paired responses by group


```{r}
reponse_table <- persona %>%
  dplyr::filter(qbase == "Q7.3" | qbase == "Q7.4") %>%
  dplyr::select((!!sym(gp_col)), id_person, qbase, response) %>%
  tidyr::pivot_wider(names_from = qbase, values_from = response) %>%
  dplyr::filter(! (is.na(Q7.3) & is.na(Q7.4))) %>%
  dplyr::arrange((!!sym(gp_col)), id_person)

reponse_table %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 1), color = "white", background = .GlobalEnv$COLORS[1]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 2), color = "white", background = .GlobalEnv$COLORS[2]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 3), color = "white", background = .GlobalEnv$COLORS[3]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 4), color = "white", background = .GlobalEnv$COLORS[4]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 5), color = "white", background = .GlobalEnv$COLORS[5])
  
```
