---
title: "Clinician Sub-Analysis"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
params:
  num_clusters: 3
  survey_data: "survey_likert"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(RColorBrewer)
library(knitr)
library(kableExtra)

source(here("./R/plot_question_bar.R"))
source(here("./R/likert.R"))
source(here("./R/recode_occupation.R"))

# params <- list(); params$num_clusters <- 4; params$survey_data <- "survey_likert"

print(params)

fs::dir_create(here("./output", "persona",
                    glue::glue("{params$survey_data}"),
                    glue::glue("clinician_subanalysis_{params$num_clusters}")), recurse = TRUE)

# generated from clustering.Rmd
clinician <- readr::read_tsv(here(glue::glue("./data/final/persona/{params$survey_data}/persona_clinician_group_{params$num_clusters}.tsv")))
metadata <- jsonlite::read_json(here("./data/final/01-surveys/01-self_assessment_persona_metadata.json"))

gp_col <- glue::glue("group{params$num_clusters}")

# i'm using !!sym because we're not in a function
counts <- clinician %>%
  dplyr::group_by((!!sym(gp_col)), qbase, question_text) %>%
  dplyr::count(response) %>%
  tidyr::drop_na() %>%
  dplyr::ungroup()
```

```{r}
knitr::include_graphics(here::here(glue::glue(
  "./output/persona/{params$survey_data}/dendogram_{params$num_clusters}.png"
)))
```

```{r}
knitr::include_graphics(here::here(glue::glue(
  "./output/persona/{params$survey_data}/dendogram_{params$num_clusters}_gg.png"
)))
```

Missing groups come from data used for validation keeping complete cases.

```{r}
clinician %>%
  dplyr::filter(qbase == "Q1.3") %>%
  dplyr::count( (!!sym(gp_col)) )
```

```{r}
plot_occupation_group <- function(dat, gp_col = NULL) {
  # user string; function quote
  ex_gp_col <- rlang::parse_expr(gp_col)

  occupation_cluster <- dat %>%
    dplyr::filter(qbase == "Q2.3") %>%
    dplyr::select(id_person, id_response, question_part, response, qbase, qid, !!ex_gp_col) %>%
    .GlobalEnv$recode_occupation(response) %>%
    tidyr::drop_na()
  
  g <- ggplot(occupation_cluster, aes(x = occupation_group,
                                      fill = as.factor(!!ex_gp_col))) +
    geom_bar(position = "dodge") +
    scale_fill_discrete(name = "Cluster") +
    coord_flip() +
    theme_minimal()

  return(g)
}
```


```{r}
g <- plot_occupation_group(clinician, gp_col)
print(g)
```

```{r}
pth <- here("./output",
            "persona",
            glue::glue("{params$survey_data}"),
            glue::glue("clinician_subanalysis_{params$num_clusters}"),
            glue::glue("group_cluster_occupation_clinician.png"))
ggsave(filename = pth, plot = g, width = 8, height = 5, units = "in")
```

# Question Bar Graphs

```{r}
for (q in unique(counts$qbase)) {
  print(q)
  #browser()
  if (q %in% c("Q7.2", "Q7.3", "Q7.4")) {next}

  g <- plot_question_bar(dplyr::filter(counts, qbase == q), fill_var = gp_col)
  pth <- here("./output",
              "persona",
              glue::glue("{params$survey_data}"),
              glue::glue("clinician_subanalysis_{params$num_clusters}"),
              glue::glue("{q}.png"))
  print(pth)
  print(g)
  ggsave(filename = pth, plot = g, width = 8, height = 5, units = "in")
}
```

# Likert table

```{r, fig.width=8, fig.height=5}
qbase <- "Q7.2"
qid <- get_qid_from_qbase(clinician, qbase)
qid

likert_q_count <- likert_calculation(clinician, metadata, qbase, qid, grp_var = gp_col)
g <- likert_plot(likert_q_count %>% tidyr::drop_na(), facet_1 = gp_col)
g
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("clinician_subanalysis_{params$num_clusters}"),
                       "likert.png"),
       plot = g,
       width = 13,
       height = 8,
       units = "in")
```

## Heatmap by proportion in group

```{r}
likert_q_count_n_prop <- likert_q_count %>%
  dplyr::mutate(rel_freq = n / sum(n),
                pct = round(rel_freq * 100))
```

```{r, fig.width=13, fig.height=8}
g <- likert_plot(likert_q_count_n_prop %>% tidyr::drop_na(), count_col = "pct", facet_1 = gp_col)
g
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("group_descriptives_{params$num_clusters}"),
                       "likert_prop.png"),
       plot = g,
       width = 13,
       height = 8,
       units = "in")
```

```{r}
# likert plot for poster
poster <- likert_plot(likert_q_count_n_prop %>% tidyr::drop_na(),
                      count_col = "pct") +
  facet_wrap(as.formula(glue::glue("~ {gp_col}")), ncol = 1) +
  theme(legend.position = "none",
        text = element_text(size=12)) +
  ggtitle("Percent of Responses per Group")
poster
```

```{r}
ggsave(filename = here("./output",
                       "persona",
                       glue::glue("{params$survey_data}"),
                       glue::glue("clinician_subanalysis_{params$num_clusters}"),
                       "likert_prop_poster.png"),
       plot = poster,
       width = 8,
       height = 13,
       units = "in")
```


# Free Response

- Q7.3: Please share what you most hope to learn from participating in this workshop and/or workshop series. 
- Q7.4: What do you want to know or be able to do after this workshop (or series of sessions) that you don't know or can't do right now?

## Paired responses by group


```{r}
reponse_table <- clinician %>%
  dplyr::filter(qbase == "Q7.3" | qbase == "Q7.4") %>%
  dplyr::select((!!sym(gp_col)), id_person, qbase, response) %>%
  tidyr::pivot_wider(names_from = qbase, values_from = response) %>%
  dplyr::filter(! (is.na(Q7.3) & is.na(Q7.4))) %>%
  dplyr::arrange((!!sym(gp_col)), id_person)

reponse_table %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 1), color = "white", background = .GlobalEnv$COLORS[1]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 2), color = "white", background = .GlobalEnv$COLORS[2]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 3), color = "white", background = .GlobalEnv$COLORS[3]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 4), color = "white", background = .GlobalEnv$COLORS[4]) %>%
  kableExtra::row_spec(which(reponse_table[gp_col] == 5), color = "white", background = .GlobalEnv$COLORS[5])
  
```
