---
title: "PCA and Factor Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(readr)
library(tibble)
library(psych)
library(here)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
numeric <- readr::read_csv(here::here("./data/final/persona/01-participant_numeric.csv"))

numeric <- tibble::column_to_rownames(numeric, var = "id")
scaled <- scale(numeric)
```

# EFA

Taken from: https://www.statmethods.net/advstats/factor.html

```{r}
# Determine Number of Factors to Extract
library(nFactors)
ev <- eigen(cor(scaled)) # get eigenvalues
ap <- parallel(subject = nrow(scaled),
               var=ncol(scaled),
               rep=100,
               cent=.05)
nS <- nScree(x=ev$values,
             aparallel=ap$eigen$qevpea)

png(here::here("./output/persona/efa_eigen_scree.png"))
plotnScree(nS)
dev.off()
```


### 2 Loadings

```{r}
efa2 <- factanal(scaled, factors = 2, rotation = "varimax")
print(efa2, digits=2, cutoff=.3, sort=TRUE)
```

```{r}
efa2 <- factanal(scaled, factors = 2, rotation = "promax")
print(efa2, digits=2, cutoff=.3, sort=TRUE)

saveRDS(efa2, here("./data/final/persona/efa2-promax.RDS"))
```

```{r}
f2 <- fa(scaled, 2)
factor.stats(scaled, f2, n.obs = nrow(scaled))

f2o <- fa(scaled, 2, fm="pa", rotate="Promax")
factor.stats(scaled, f2o, n.obs = nrow(scaled))
```


```{r}
load <- efa2$loadings[, 1:2]
plot(load,type="n") # set up plot
text(load,labels=names(numeric),cex=.7) # add variable names 
```

### 3 Loadings

```{r}
efa3 <- factanal(scaled, factors = 3, rotation = "varimax")
print(efa3, digits=2, cutoff=.3, sort=TRUE)
```

```{r}
efa3 <- factanal(scaled, factors = 3, rotation = "promax")
print(efa3, digits=2, cutoff=.3, sort=TRUE)
```

```{r}
saveRDS(efa3, here("./data/final/persona/efa3-promax.RDS"))
```

```{r}
psych::fac(scaled, nfactors = 3, rotate = "promax")
```

```{r}
psych::fac(scaled, nfactors = 2, rotate = "promax")
```




```{r}
f3 <- fa(scaled, 3)
factor.stats(scaled, f3, n.obs = nrow(scaled))

f3o <- fa(scaled, 3, fm="pa", rotate="Promax")
factor.stats(scaled, f3o, n.obs = nrow(scaled))
```


```{r}
load <- efa3$loadings[, 1:2]
plot(load,type="n") # set up plot
text(load,labels=names(numeric),cex=.7) # add variable names 
```

### 4 Loadings

```{r}
efa4 <- factanal(scaled, factors = 4, rotation = "varimax")
print(efa4, digits=2, cutoff=.3, sort=TRUE)
```

```{r}
efa4 <- factanal(scaled, factors = 4, rotation = "promax")
print(efa4, digits=2, cutoff=.3, sort=TRUE)
```

```{r}
load <- efa3$loadings[, 1:2]
plot(load,type="n") # set up plot
text(load,labels=names(numeric),cex=.7) # add variable names 
```

Looking at the CFA results and interprebility, I'm using 3 factors.
Results using promax rotation.

- Programming:
  - Q3.3: How familiar are you with interactive programming languages like Python or R?
- Stats:
  - Q6.2: If you were given a dateset containing an individual's smoking status (binary variable) and whether or not they have hypertension (binary variable), would you know how to conduct a statistical analysis to see if smoking has an increased relative risk or odds of hypertension? Any type of model will suffice.
- Data:
  - Q4.4: Do you know what "long" and "wide" data are?

```{r}
scaled_subq <- scaled[, c("Q3.3", "Q6.2", "Q4.4")]
```

```{r}
scaled_subq
```

```{r}
dist <- dist(scaled_subq, method = "euclidean")
hc_ward <- hclust(dist, method = "ward.D2")
# Plot the result
plot(hc_ward, cex = 0.6, hang = -1)
```

```{r}
sub_grp4 <- cutree(hc_ward, k = 4)
sub_grp5 <- cutree(hc_ward, k = 5)
```

```{r}
numeric <- numeric %>%
  dplyr::mutate(
    group4 = sub_grp4,
    group5 = sub_grp5,
  )
rownames(numeric) <- rownames(scaled_subq)
```

```{r}
#png(here::here("./output/persona/dendogram_4.png"), width = 13, height = 8, units = "in", res = 72)
plot(hc_ward, cex = 0.6)
rect.hclust(hc_ward, k = 4, border = 2:5)
#dev.off()
```

```{r}
#png(here::here("./output/persona/dendogram_5.png"), width = 13, height = 8, units = "in", res = 72)
plot(hc_ward, cex = 0.6)
rect.hclust(hc_ward, k = 5, border = 2:6)
#dev.off()
```

Comparing original dendogram results with subsetted results

```{r}
persona4 <- readr::read_tsv(here("./data/final/persona/persona_group_4.tsv"))

id4 <- persona4 %>%
  dplyr::select(id, group4) %>%
  dplyr::distinct() %>%
  dplyr::arrange(group4) %>%
  dplyr::mutate(id = as.character(id))
```

```{r}
efa4 <- numeric
efa4$id <- rownames(efa4)
efa4 <- efa4[, c("id", "group4")] %>%
  dplyr::arrange(group4) %>%
  #dplyr::mutate(id = as.numeric(id)) %>%
  {.}
```

```{r}
id4
```

```{r}
efa4
```

```{r}
joined_id_group <- dplyr::left_join(id4, efa4, by = "id", suffix = c("_dendo", "_efa")) %>%
  dplyr::mutate(
    group_match = dplyr::if_else(group4_dendo == group4_efa, TRUE, FALSE)
  )
```

```{r}
joined_id_group %>%
  dplyr::group_by(group4_dendo) %>%
  dplyr::summarize(matches = sum(group_match),
                   total = n(),
                   prop_match = round(matches / total, digits = 2)) %>%
  knitr::kable()
```


# CART

- 6.2
  1. "I wouldn't know where to start",
  2. "I could struggle through, but not confident I could do it",
  3. "I could struggle through by trial and error with a lot of web searches",
  4. "I could do it quickly with little or no use of external help"

- 3.3
  1. "I do not know what those are",
  2. "I have heard of them but have never used them before",
  3. "I have installed it, but have only done simple examples with them",
  4. "I have written a small program with them before",
  5. "I use it to automate certain repetitive tasks",
  6. "I have small side projects that I program in it",
  7. "I program in them for work"

- 4.4
  1. "I have never heard of the term",
  2. "I have heard of it but don't remember what it is.",
  3. "I have some idea of what it is, but am not too clear",
  4. "I know what it is and could explain what it pertains to"

```{r}
library(rpart)

tree_dat <- numeric %>%
  dplyr::select(group4, starts_with("Q")) %>%
  tibble::rownames_to_column("id") %>%
  dplyr::left_join(efa4, by="id", suffix = c("_dendo", "_efa")) %>%
  dplyr::select(group4_efa, starts_with("Q"))

model <- rpart(group4_efa ~ ., data = tree_dat)

png(here("./output/persona/cart.png"))
par(xpd = NA) # otherwise on some devices the text is clipped
plot(model)
text(model, digits = 3)
dev.off()
```

```{r}
library(rpart)
tree_dat_all_q <- numeric %>%
  select(group4, starts_with("Q"))

model_all_q <- rpart(group4 ~ ., data = tree_dat_all_q)

par(xpd = model_all_q) # otherwise on some devices the text is clipped
plot(model_all_q)
text(model_all_q, digits = 3)
```

# CFA

Using the 3 factor EFA results:

```{r}
print(efa3, digits=2, cutoff=.3, sort=TRUE)
```


```{r}
ds_model <- '
programming =~ Q3.1 + Q3.3 + Q3.4 + Q3.5 + Q5.2
statistics =~ Q6.1 + Q6.2 + Q6.3
data =~ Q4.4 + Q6.4
'
```

```{r}
library(lavaan)
library(semPlot)
```

```{r}
cfa <- lavaan::cfa(ds_model, data = scaled)
summary(cfa, standardized = TRUE)
```

```{r}
semPlot::semPaths(cfa)
```

```{r}
png(here::here("./output/persona/cfa.png"),
    width = 8, height = 5, units = "in", res = 300)
semPlot::semPaths(cfa, "std")
dev.off()
```

## CFA Fit

```{r}
fitmeasures(cfa, fit.measures = "all")
```

```{r}
fitmeasures(cfa, c("cfi", "tli", "rmsea", "srmr", "wrmr"))
```

This looks decent, but the rmsea is above 0.05.


