---
title: "CART"
author: "Daniel Chen"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# CART

- 6.2
  1. "I wouldn't know where to start",
  2. "I could struggle through, but not confident I could do it",
  3. "I could struggle through by trial and error with a lot of web searches",
  4. "I could do it quickly with little or no use of external help"

- 3.3
  1. "I do not know what those are",
  2. "I have heard of them but have never used them before",
  3. "I have installed it, but have only done simple examples with them",
  4. "I have written a small program with them before",
  5. "I use it to automate certain repetitive tasks",
  6. "I have small side projects that I program in it",
  7. "I program in them for work"

- 4.4
  1. "I have never heard of the term",
  2. "I have heard of it but don't remember what it is.",
  3. "I have some idea of what it is, but am not too clear",
  4. "I know what it is and could explain what it pertains to"

```{r}
library(rpart)

tree_dat <- numeric %>%
  dplyr::select(group4, starts_with("Q")) %>%
  tibble::rownames_to_column("id") %>%
  dplyr::left_join(efa4, by="id", suffix = c("_dendo", "_efa")) %>%
  dplyr::select(group4_efa, starts_with("Q"))

model <- rpart(group4_efa ~ ., data = tree_dat)

png(here("./output/persona/cart.png"))
par(xpd = NA) # otherwise on some devices the text is clipped
plot(model)
text(model, digits = 3)
dev.off()
```

```{r}
library(rpart)
tree_dat_all_q <- numeric %>%
  select(group4, starts_with("Q"))

model_all_q <- rpart(group4 ~ ., data = tree_dat_all_q)

par(xpd = model_all_q) # otherwise on some devices the text is clipped
plot(model_all_q)
text(model_all_q, digits = 3)
```

# CFA

Using the 3 factor EFA results:

```{r}
print(efa3, digits=2, cutoff=.3, sort=TRUE)
```


```{r}
ds_model <- '
programming =~ Q3.1 + Q3.3 + Q3.4 + Q3.5 + Q5.2
statistics =~ Q6.1 + Q6.2 + Q6.3
data =~ Q4.4 + Q6.4
'
```

```{r}
library(lavaan)
library(semPlot)
```

```{r}
cfa <- lavaan::cfa(ds_model, data = scaled)
summary(cfa, standardized = TRUE)
```

```{r}
semPlot::semPaths(cfa)
```

```{r}
png(here::here("./output/persona/cfa.png"),
    width = 8, height = 5, units = "in", res = 300)
semPlot::semPaths(cfa, "std")
dev.off()
```
